Uncovering underlying musical content in the temporal domain \\ \small Leveraging deep learning, inductive bias, and aural music skills to learn embedding spaces from raw waveforms with applications to MIR downstream tasks

### 2023-05-06 figures/neural networks/activation.tex
Updates from Overleaf

% NEURAL NETWORK activation
% https://www.youtube.com/watch?v=aircAruvnKk&list=PLZHQObOWTQDNU6R1_67000Dx_ZCJB-3pi&index=1
\begin{tikzpicture}[x=2.7cm,y=1.6cm]
  \message{^^JNeural network activation}
  \def\NI{5} % number of nodes in input layers
  \def\NO{4} % number of nodes in output layers
  \def\yshift{0.4} % shift last node for dots

@@

\end{tikzpicture}

>>>>

\begin{figure}
    \centering
    \scalebox{0.8}{
% NEURAL NETWORK activation
% https://www.youtube.com/watch?v=aircAruvnKk&list=PLZHQObOWTQDNU6R1_67000Dx_ZCJB-3pi&index=1
\begin{tikzpicture}[x=2.7cm,y=1.6cm]
\centering
  \message{^^JNeural network activation}
  \def\NI{5} % number of nodes in input layers
  \def\NO{4} % number of nodes in output layers
  \def\yshift{0.4} % shift last node for dots

@@

\end{tikzpicture}
}
    \caption[Activation function]{\small{The input layer has "n" nodes, represented as $a_i^{(0)}$, where $i$ ranges from 1 to n. The output layer has "m" nodes, represented as $a_j^{(1)}$, where $j$ ranges from 1 to m. The connections between the input and output layers are defined by the weight matrix $\mathbf{W}^{(0)} \in \mathbb{R}^{m\times n}$, with each element represented as $w_{j,i}$, and the bias vector $\mathbf{b}^{(0)} \in \mathbb{R}^{m}$, with each element represented as $b_j^{(0)}$.The activation function, \\sigma, is applied element-wise to the product of the weight matrix and the input vector plus the bias vector}}

    \label{fig:activation_function}

\end{figure}

### 2023-05-06 figures/neural networks/colour CNN.tex
Updates from Overleaf

\end{tikzpicture}
    \caption{High-level illustration of a CNN}
    \label{fig: CNN colour diagram}
\end{figure}

>>>>

\end{tikzpicture}
    \caption[High-level CNN diagram.]{\small{High-level illustration of a CNN. It consists of an input layer that takes in raw data, followed by a downsampling process using convolution and pooling layers to extract features. Finally, fully-connected hidden layers are added for learning the global structure, and an output layer produces the desired predictions.}}
    \label{fig: CNN colour diagram}
\end{figure}

### 2023-05-06 figures/neural networks/convolution.tex
Updates from Overleaf

    \caption{\small{Illustration of the 2D convolution operation on a binary input matrix $\mathbf I$ with a kernel matrix $\mathbf K$. The orange-highlighted region in $\mathbf I$ represents the portion convolved with the kernel, resulting in the blue-highlighted element in the output matrix $\mathbf{I * K}$.}}
    \label{fig: Convolution}
\end{figure}

>>>>

    \caption[Illustration of the 2D convolution operation.]{\small{Illustration of the 2D convolution operation on a binary input matrix $\mathbf I$ with a kernel matrix $\mathbf K$. The orange-highlighted region in $\mathbf I$ represents the portion convolved with the kernel, resulting in the blue-highlighted element in the output matrix $\mathbf{I * K}$.}}
    \label{fig: Convolution}
\end{figure}

### 2023-05-06 introduction/Subsections/Assumptions.tex
Updates from Overleaf

Abstract features, such as melody, harmony, rhythm, tempo, form, and expression, can be identified irrespective of waveform or production style. This approach offers objective and comprehensive insight into a piece's musical content and meaning, analogous to how analyzing sheet music unveils a composition's underlying structure and intent.

@@

>>>>

Abstract features, such as melody, harmony, rhythm, tempo, form, and expression, can be identified irrespective of waveform or production style. This approach offers an objective and comprehensive insight into a piece's musical content and meaning, analogous to how analyzing sheet music unveils a composition's underlying structure and intent.

@@

\subsection{Beyond Digital Signal Processing (DSP)}

We argue that MIR research needs to incorporate a more balanced approach that considers the interdisciplinary nature of music and the importance of other domains beyond DSP.

### 2023-05-06 introduction/Subsections/Introduction.tex
Updates from Overleaf

Music\footnote{In this thesis, "music" refers to the Western tradition, characterized by conventions, practices, and aesthetics that have developed primarily in Europe and North America. This usage is not intended to deny the diversity and richness of other musical traditions worldwide but instead reflects the present study's focus on a specific cultural context.}, an essential component of human culture, history, and society, has existed since ancient times. It reflects the changes in our cultures, traditions, and beliefs as it evolves. Studying music, however, has always been a challenging and complex task due to its subjective nature, relative and mutable ground truth, and the sheer variety of musical styles, genres, and worldwide cultures.

>>>>

Music \footnote{In this thesis, "music" refers to the Western tradition, characterized by conventions, practices, and aesthetics that have developed primarily in Europe and North America. This usage is not intended to deny the diversity and richness of other musical traditions worldwide but instead reflects the present study's focus on a specific cultural context.}, an essential component of human culture, history, and society, has existed since ancient times. It reflects the changes in our cultures, traditions, and beliefs as it evolves. Studying music, however, has always been a challenging and complex task due to its subjective nature, relative and mutable ground truth, and the sheer variety of musical styles, genres, and worldwide cultures.

### 2023-05-06 introduction/Subsections/Motivation.tex
Updates from Overleaf

\begin{figure}[ht]
\includegraphics[clip,width=\columnwidth]{figures/schenkerian analysis/SchubertOp4no3.png}% 
\caption{\small{Small excerpt of \textit{Wandrers Nachtlied, Op. 4, D. 224} by Franz Schubert. We can see the passage's original score, the schenkerian unfolding of the melody, the chord degrees analysis, and the tonal function.}}
\label{fig:Wandrers Nachtlied, Op. 4, D. 224}
\end{figure}

>>>>

\begin{figure}[ht]
\includegraphics[clip,width=\columnwidth]{figures/schenkerian analysis/SchubertOp4no3.png}% 
\caption[Small excerpt of \textit{Wandrers Nachtlied, Op. 4, D. 224} by Franz Schubert.]{\small{Small excerpt of \textit{Wandrers Nachtlied, Op. 4, D. 224} by Franz Schubert. Passage's original score, the schenkerian unfolding of the melody, the chord degrees analysis, and their tonal function.}}
\label{fig:Wandrers Nachtlied, Op. 4, D. 224}
\end{figure}

### 2023-05-06 methods/subsections/model architecture and training strategy.tex
Updates from Overleaf

In machine learning and pattern recognition, invariance refers to the property of a model or algorithm being robust to certain transformations or changes in the input data. In other words, an invariant model should produce the same or similar output for different instances of the same object, even if they have undergone some changes.

@@

In summary, known invariances are explicitly incorporated into the design of a model or algorithm to improve its robustness and generalization. In contrast, the model learns unknown invariances from the data without explicit instruction.

\begin{equation}
\forall x \in X, \forall y \in Y, \forall \rho \in \mathcal{T}, f(x) = f(y) \implies f(\rho(x)) = f(\rho(y))
\end{equation}

In this expression, $X$ and $Y$ represent sets of input data, $\rho$ represents a transformation function, and $f$ represents a machine learning model. The expression states that for all inputs $x$ and $y$ in $X$ and $Y$, if $f(x) = f(y)$, then $f(\rho(x)) = f(\rho(y))$.

@@

Contrastive Learning of Musical Representations (CLMR) is an approach that combines the principles of Self-Supervised Learning (SSL) and Contrastive Learning of Musical Representations (CLMR) \cite{CLMR2021} to learn discriminative and valuable representations of music without relying on explicit labels. By leveraging the structure and content of music and contrasting different augmentations or transformations of the same musical piece against randomly (or not) selected pieces, CSSLMR models can effectively capture the underlying structure and known invariance in the data.

@@

$y_{c_o}(n)$ is the output for the $c_o$-th output channel at position $n$
$C_{\text{in}}$ is the number of input channels
$x_{c_i}(n)$ is the input for the $c_i$-th input channel at position $n$
$w_{c_o, c_i}(k)$ is the kernel weight for the $c_o$-th output channel and $c_i$-th input channel at position $k$
$K$ is the kernel size
The stride is set to 1 and padding is set to 1 (equal padding $\lfloor K/2 \rfloor$ on both sides)
s
    \item Batch normalization. We process a batch of data containing anchor, positive, and negative samples. We pad the waveforms to the same  maximum length among batch, performs online triplet mining to find the hardest negatives, and returns the padded anchors, positives, and hardest negatives.

@@

The $\max(0, ...)$ function computes the maximum of the expression within the parentheses and 0. This function ensures that the loss is only calculated for triplets that violate the margin constraint, where the negative sample is closer to the anchor than the positive sample by at least the margin. The loss is zero when the distance between the positive and negative samples is greater than the margin. Otherwise, the loss is proportional to the extent to which the constraint is violated.

>>>>

In pattern recognition, invariance refers to the property of a model or algorithm being robust to certain transformations or changes in the input data. In other words, an invariant model should produce the same or similar output for different instances of the same object, even if they have undergone some changes.

@@

\begin{equation}
\label{eq:invariance_to_particular_transformation_function}
\forall x \in X, \forall y \in Y, \forall \rho \in \mathcal{T}, f(x) = f(y) \implies f(\rho(x)) = f(\rho(y))
\end{equation}

In this expression \ref{eq:invariance_to_particular_transformation_function}, $X$ and $Y$ represent sets of input data, $\rho$ represents a transformation function, and $f$ represents a machine learning model. The expression states that for all inputs $x$ and $y$ in $X$ and $Y$, if $f(x) = f(y)$, then $f(\rho(x)) = f(\rho(y))$.

@@

Contrastive Learning of Musical Representations (CLMR) \cite{CLMR2021} aims to learn discriminative and valuable representations of music without relying on explicit labels. By leveraging the structure and content of music and contrasting different augmentations or transformations of the same musical piece against randomly generated pieces, CLMR models can effectively capture the underlying known invariances that are explicitly incorporated into the design of a model to improve its robustness and generalization.

In contrast, the model learns unknown invariances from the data without explicit instruction.

@@

$y_{c_o}(n)$ is the output for the $c_o$-th output channel at position $n$
$C_{\text{in}}$ is the number of input channels
$x_{c_i}(n)$ is the input for the $c_i$-th input channel at position $n$
$w_{c_o, c_i}(k)$ is the kernel weight for the $c_o$-th output channel and $c_i$-th input channel at position $k$
$K$ is the kernel size
The stride is set to 1, and padding is set to 1 (equal padding $\lfloor K/2 \rfloor$ on both sides)
s
    \item Batch normalization. We process a batch of data containing anchor, positive, and negative samples. We pad the waveforms to the same  maximum length among the collection, perform online triplet mining to find the hardest negatives and return the padded anchors, positives, and hardest negatives.

@@

The $\max(0, ...)$ function computes the maximum of the expression within the parentheses and 0. This function ensures that the loss is only calculated for triplets that violate the margin constraint, where the negative sample is closer to the anchor than the positive sample by at least the margin. The loss is zero when the distance between the positive and negative samples exceeds the margin. Otherwise, the loss is proportional to the extent to which the constraint is violated.
